Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-23 src/tex/ms.tex
Update

\begin{abstract}
	We investigate the possibility of improving the accuracy of the
	phenomenological waveform model, IMRPhenomD, by jointly optimizing all the
	calibration coefficients at once, given a set of numerical relativity (NR)
	waveforms. When IMRPhenomD was first calibrated to NR waveforms, different
	parts (i.e., the inspiral, merger, and ringdown) of the waveform were calibrated separately. 
	Using \ripple, a library of waveform models compatible with automatic differentiation, we can, for the first time,
	perform gradient-based optimization on all the waveform coefficients at the same time.
	This joint optimization process allows us to capture previously ignored correlations between separate parts of the waveform.
	We found that after recalibration, the median mismatch between the model and NR waveforms decreases by $50\%$.
	We further explore how different regions of the source parameter space respond
	to the optimization procedure. We find that the degree of improvement
	correlates with the spins of the source. This work shows a promising avenue
	to help understand and treat systematic error in waveform models.
\end{abstract}

@@

Many data analysis tasks in gravitational wave (GW) astrophysics, such
as match filtering~\citep{owen1996search, owen1999matched} and parameter 
estimation~\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
models. Since the generation of numerical relativity (NR) waveforms is
prohibitively expensive, the community has constructed waveform approximants
that can be evaluated much faster. There are three families of commonly used GW
approximants: the effective-one-body (EOB) \citep{ossokine2020multipolar,
cotesta2020frequency, taracchini2014effective}, NR
surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2},
and phenomenological (Phenom) models \citep{husa2016frequency,
khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
detailed construction of each model is different, they all have a set of
internal parameters that can be calibrated to NR waveforms. The quality of the
waveform model is therefore determined by the ansatz used and the accuracy of the
calibrated parameters.

@@

The LIGO-VIRGO-KAGRA (LVK) collaboration
~\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
VIRGO:2014yos,KAGRA:2020tym} recently started their fourth observational
run on May 26, 2023. Impressively, they are
expected to double the total number of observed binary black holes (BBHs)~\citep{abbott2020prospects}.
Morover, the improved sensitivity also implies that we expect to
detect individual events with a higher signal-to-noise ratio (SNR) than ever before. This means
we can resolve more features in the signal, therefore putting more
stringent requirements on the accuracy of our waveform model \citep{purrer2020gravitational, hu2022assessing}.

@@

Because of the large number of calibration parameters (often a few hundred if not
more), waveform models are usually calibrated separately for the inspiral, merger,
and ringdown parts of the waveform \citep{khan2016frequency, 
santamaria2010matching, pratten2021computationally}. This
ignores the correlation between different parts of the waveform model and limits
it quality. Recently, there has been an effort to rebuild waveform models
\citep{khan2016frequency} using programming languages that support automatic differentiation (AD) 
\citep{ripple, Iacovelli:2022bbs, Iacovelli:2022mbg, Coogan:2022qxs}; AD is a technique 
used to compute machine precision derivatives of functions 
without the issues of scaling up to high dimension or expression swelling. In
particular, \ripple~\citep{ripple} exposes the calibration parameters to the
user. This allows us to make use of common techniques from
machine learning, such as gradient descent and back propagation \citep{jax2018github, 
pytorch, tensorflow2015}, to improve the calibration of the waveform models.

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of a waveform model, IMRPhenomD \citep{khan2016frequency, husa2016frequency}, by jointly optimizing all the
calibration coefficients for given a set of NR waveforms. Using a similar set of NR
waveforms to those used in \citep{khan2016frequency, husa2016frequency},
we demonstrate that one can improve the match between IMRPhenomD and NR
waveforms over a decently sized parameter space, up to a mass ratio $q=8$. We additionally
explore how different parts of the source parameter space (e.g. the primary and
second spins) respond to the optimization procedure by optimizing the waveform
separately for different regions. This can help in understanding
whether the waveform model ansatz performs equally well in different regions of the
parameter space.

The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
review the parameterization of the IMRPhenomD model and the mismatch function
that is used as an objective/loss function for the calibration, followed by outlining the
specific optimization scheme used for recalibration. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of the optimized waveforms with the original waveforms. We also show how the
optimization result differs as a function of the source parameters. In
Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
procedure and the procedure used in \citep{khan2016frequency}. We also explain how the reduced spin
parameterization affects the accuracy of the model.

@@

In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
\end{align}
where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively.

@@

%\kw{Need to work on this paragraph}
Although initially independent, the stitching procedure means that each section
of the waveform intrinsically depends on the full set of $\lambda$'s. 
A slightly inaccurate set of $\lambda$'s can therefore lead to inaccuracies in
the generated waveforms. 
Thus, the calibration of these coefficients is crucial to the accuracy
of IMRPhenom GW models. 
Importantly, since the fitting was performed on the individual segments,
the final waveform is not guaranteed to have $\lambda$'s close to global minima.

@@

At the time of construction this piece-wise approach was necessary since
$\lambda$ has 209 components, making the fitting to NR simulations computationally prohibitive.
Here, for the first time we recalibrate the $\lambda$ coefficients jointly. 
This is made possible by the use of gradient-based optimization algorithms,
enabled by AD from \jax\, and {\ripple}, which are significantly more efficient in high dimensions.
% , gradients of IMRPhenomD can be easily obtained, thus
% allowing the use of gradient-based algorithms for us to recalibrate the model.

@@

In order to optimize the coefficients, we need to define a loss function that
quantifies the difference between the waveform model and the target NR simulations which we want to match.
Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm,husa2016frequency}. 
It is defined as
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ are the two GW waveforms we are comparing, and $t_0$ and $\phi_0$
are a relative time and phase shift respectively. 
The inner product, $\langle h_1, h_2 \rangle$, is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
$S_n(f)$ is the power spectral density (PSD), and $f_{\mathrm{max}}$ ($f_{\mathrm{min}}$) is
the maximum (minimum) frequency for the integration.
We note here that the mismatch can be viewed as the mean square error (MSE) between the
two waveforms.

Since we wish to optimize the model over the whole parameter space,
we need to compare multiple model generated waveforms with NR waveforms.
However, the mismatch is only defined for two input waveforms at a particular set of intrinsic parameters.
We therefore define the loss function as an average of training waveforms in two ways,
the simple average of mismatches and the normalized average of mismatches,  
\begin{align}
	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
	\mathcal{L}_{\mathrm{nl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
\end{align}	
where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
training waveform, and $N$ is the total number of training waveforms.

The two loss functions are chosen to give different 
behaviour during the optimization process.
The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
but is prone to be dominated by a single point in parameter space with a large mismatch.
Other points with smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such a loss function.
Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
by encouraging the waveform to improve at each training point at a similar rate.
The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
and preferencially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
In this paper, we show the results of using both loss functions 
and examine the differences between them.

@@

To compute the loss functions, we have to choose NR waveforms for calculating the mismatch.
Originally, 19 NR simulations  were used to calibrate IMRPhenomD \citep{khan2016frequency}; 9 of these 
are from the SXS catalog \citep{boyle2019sxs} and 10 BAM simulations. 
As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
Instead, we take the same nine waveforms from the SXS catalog plus two additional waveforms which closely mimic
two of the low mass ratio BAM simulations. 
The remaining BAM simulations have no close SXS counterparts, and are therefore not included in the training set.
Our main results therefore uses 11 NR waveforms for calibration which are listed in Tab.~\ref{tab:q148}.
Additional NR waveforms which are used for further calibration are listed in Tab.~\ref{tab:q1248}.

The training set has a maximum mass ratio of eight due to the lack 
of high mass ratio simulations in the SXS catalog.
In fact, the SXS catalog only has NR waveforms with $q\leq10$.
Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$,
as most BBH events observed by LIGO and Virgo have $q\leq8$.

The SXS NR waveforms are given as time-series strain.
Since IMRPhenomD is modeled in the frequency domain, we need to Fourier transform the simulation results
in order to compute the mismatch, \eqref{eq:mismatch}.
For this, we taper the time-series using Tukey window~\citep{tukey}\footnote{
	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
} before using standard FFT routines to compute the Fourier transform.

In addition to choosing the NR waveforms for the training set,
one needs to choose a relevant PSD for mismatch.
We have opted to use a flat PSD for the mismatch calculation,
as it provides results that are independent of the detector sensitivity and mass scale.
The use of a flat PSD ensures that the improvement in accuracy is due mainly
to the difference in high-dimensional fitting.
Additionally, we are interested in examining the effect of introducing a detector PSD
on the optimization process. For this, we have chosen the
zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
Since the total mass of the system scales with the frequency of the waveform,
we must choose a corresponding mass scale to match the frequency range of our noise PSD.
To demonstrate the effect of introducing a detector specific PSD,
we selected an arbitrary mass scale of $M=50M_{\odot}$, as binaries
of this mass are commonly observed by the LIGO and Virgo detectors~\citep{gwtc1, gwtc2, gwtc21, gwtc3}.
%\te{This choice seems pretty arbitrary. Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70. This choice is important and should be more informed.}We will then examine the differences in optimization.

@@

We point out that our treatment of NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
In the original calibration process, the training waveforms are hybrids of NR and SpinAlignedEOB (SEOB) waveforms.
The low frequency inspiral part is taken from the SEOB
waveforms while the rest is taken from NR simulations.
Instead, we solely use NR waveforms for calibration since most NR waveforms used (for both training and testing)
have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, to contain part of the inspiral segment and all
merger and ringdown frequency information. We use the frequency limits
$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
the IMRPhenomD's frequency range, except the minimum frequency is set slightly higher
than in the original calibration due to the NR simulation length.
When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum
frequency \citep{santamaria2010matching}.
We use the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
which is sufficient to capture all features of the GW strain.

With the loss function evaluated, we apply gradient descent to optimize the tunable
coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
original coefficients given in \citep{khan2016frequency} because they likely lie in the
neighborhood of the minimum that we wish to find. 
We fix our learning rate, $\alpha$, to be $=10^{-6}$, which is small enough to ensure we 
don't move far from the minimum.
Finally, we stop the optimization when the validation loss stops decreasing.
This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.

@@

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio
	$q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed
	here, 9 of them are also used in the original IMRPhenomD calibration.
	\citep{khan2016frequency}}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{nl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.

>>>>

\begin{abstract}
	We investigate the possibility of improving the accuracy of a
	phenomenological waveform model, IMRPhenomD, by jointly optimizing all the
	calibration coefficients at once, given a set of numerical relativity (NR)
	waveforms. When IMRPhenomD was first calibrated to NR waveforms, different
	parts of the waveform were calibrated separately. Using \ripple, a library
	of waveform models compatible with automatic differentiation, we can perform
	gradients-based optimization to all the waveform coefficients the same time,
	which should improve the quality of waveform by capturing correlations
	between previous separately fitted parts. We found that after recalibration,
	the median mismatch between the model and NR waveforms decrease by $50\%$.
	We further explore how different parts of the source parameter space respond
	to the optimization procedure. We find that the degree of improvement
	correlates with the spins of the source. This work shows a promising avenue
	to help understand and treat systematic error in waveform models.
\end{abstract}

@@

Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
as match filtering search \citep{fill} and parameter estimation\citep{fill}, rely on accurate waveform
models. Because using numerical relativity (NR) waveforms in these tasks is
prohibitively expensive, the community has constructed waveform approximants
that can be evaluated much faster. There are three families of commonly used GW
approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
and phenomenological (Phenom) models \citep{husa2016frequency,
khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
detail of construction of these models are different, they all have a set of
internal parameters that can be calibrated to NR waveforms. The quality of a
waveform model is determined by the ansatz used and the accuracy of the
calibrated parameters.

@@

The LIGO-VIRGO-KAGRA (LVK) collaboration
(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
run on May 26, 2023. Because of the sensitivity improvement, the new run is
expected to double the size of current binary black hole (BBH) observations
\citep{abbott2020prospects}. The improved sensitivity also implies we expect to
detect individual events with a higher signal-to-noise ratio (SNR). This means
we can resolve more features in the signal, at the same time put a more
stringent requirement on the accuracy of our waveform model\citep{fill}.

@@

Because of the large number of calibration parameters (often in a couple hundreds if not
more), waveform models is usually calibrated in pieces \citep{fill}. This
ignores the correlation between different parts of the waveform model and limits
it quality. Recently, there has been efforts in rebuilding waveform models
\citep{fill} that supports automatic differentiation (AD)\citep{fill}, which is
a technique used to compute derivatives of functions down to machine precision
without issues of scaling up to high dimension or expression swelling. In
particular, {\ripple} \citep{ripple} exposes the calibration parameters to the
user. This allows us to make use of infrastructures that are heavily used in
machine learning, such as gradient descent and back propagation\cite{fill}, to
improve the calibration of the waveform models.

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of a waveform model, IMRPhenomD\citep{fill}, by jointly optimizing all the
calibration coefficients given a set of NR waveforms. Training on a few NR
waveforms, we demonstrate one can improve the match between IMRPhenomD and NR
waveforms over a decently sized parameter space, up to mass ratio $q=8$ We also
explore how different parts of the source parameter space (e.g. the primary and
second spins) respond to the optimization procedure, by optimizing the waveform
separately for different regions of the parameter space. This gives us hints on
whether the waveform model performs equally well in different regions of the
parameter space.

The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
review the parameterization of the IMRPhenomD model and the mismatch function
that is used as an objective function for the calibration, followed by outlining the
specific optimization scheme used for recalibration. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of optimized waveforms with original waveforms. We also show how the
optimization result differs with waveforms of different source parameters. In
Sec.~\ref{sec:discussion}, we address the difference between our calibrating
procedure with \citep{khan2016frequency}. We also explain how reduced spin
parameterization affects the accuracy of the model.

@@

In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRhenomD uses the polynomial function:
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
\end{align}
where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively.

@@

%\kw{Need to work on this paragraph}
Although initially independent, the stiching procedure means that each section
of the waveform intrinsically depends on the full set of $\lambda$'s. 
A slightly inaccurate set of $\lambda$'s can therefore lead to inaccuracies in
the generated waveforms. 
Thus, the calibration of these coefficients is crucial to the accuracy
of IMRPhenom GW models. 
Importantly, since the fitting was performed on the individual segments,
the final waveform is not guarenteed to have $\lambda$'s close to global minima.

@@

At the time of construction this piece-wise approach was necessary since
$\lambda$ has 209 components, making the fitting to NR simulations computationally prohibitive.
Here, for the first time we recalibrate the $\lambda$ coefficients jointly. 
This is made possible by the use of gradient-based optimization algorithms,
enabled by AD from \jax\, and {\ripple}, which are significatnly more efficient in high dimensions.
% , gradients of IMRPhenomD can be easily obtained, thus
% allowing the use of gradient-based algorithms for us to recalibrate the model.

@@

In order to optimize the coefficients, we need to define a loss function that
quantifies the difference between waveform model and the target NR simulations which we want to match.
Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{husa2016frequency}. 
It is defined as
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ are the two GW waveforms we are comparing, and $t_0$ and $\phi_0$
are a relative time and phase shift respectively. 
The inner product, $\langle h_1, h_2 \rangle$, is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
$S_n(f)$ is the power spectral density (PSD), $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
the relevant maximum and minimum frequencies for the integration.
We note here that the mismatch can be viewed as the mean square error (MSE) between the
two waveforms.

\te{Up to here}

Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms. However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters. To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space. We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,  
\begin{align}
	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
\end{align}	
where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
training waveform, and $N$ is the total number of individual training waveforms.
Note that we choose to use two different averages, since they have different 
preferences in optimization base on waveform mismatches. For the first choice, 
simple average serves as the simplest choice of loss function, but is prone to 
be dominated by a single waveform with a large mismatch. Other waveforms with 
smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such optimization. Alternatively, the second choice, normalized 
average eliminates the aforementioned issue. Nevertheless, it excludes the
information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
training waveform to decrease at similar rates, hence it is hard to obtain
optimized waveforms with mismatches in the same order of magnitude. Instead,
their ratios in mismatches would remain approximately the same. Conversely,
$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
and individual mismatches would be in a similar order of magnitude after 
optimization. In this paper, we showcase the results of using both loss functions 
and examine the differences between them.

@@

To compute the loss functions, we have to take NR waveforms for calculating the mismatch. We choose 11-16 NR waveforms from the set of waveforms used in the original
calibration process as training waveforms. Originally, 19 waveforms are taken from 
NR simulations for calibrating IMRPhenomD \citep{khan2016frequency}, 
which are waveforms from the SXS catalog \citep{boyle2019sxs} or BAM simulation. 
As BAM waveforms are not publicly available, we cannot take the identical training 
set as them. Instead, we take the available waveforms from the SXS catalog to construct our loss function. Training waveforms used are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. The training waveforms chosen has maximum mass ratio to be 8. This is because SXS catalog does not have NR waveforms with extremely high mass ratio. In fact, the SXS catalog only has NR waveforms with $q\leq10$. Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed from LIGO have $q\leq8$. Hence, we calibrate IMRPhenomD with waveforms of $q\leq8$.

In the SXS catalog, NR waveforms are in the form of time-series strain. Since time-series data is oscillatory, performing optimization in the time-domain is not ideal. Hence, we transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
same intrinsic parameters. We taper the time-series using Tukey window.
\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
	entire GW strain. } 
Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series.

%\te{I think this discussion needs to be much more precise. Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper. So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD. This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
Other than NR waveforms, one need to choose a relevant noise PSD for 
mismatch. We have opted to use a flat PSD for this purpose, as it provides results that are independent of the detector sensitivity and mass scale. The use of a flat PSD ensures that the improvement in accuracy is due mainly to the difference in high-dimensional fitting and piece-wise fitting, but not due to the use of a different mass scale. Furthermore, we are interested in examining the effect of introducing a detector PSD on the optimization process. For this, we have chosen the zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}. Since the total mass of the system scales with the frequency of the waveform, we must choose a corresponding mass scale to match the frequency range of our noise PSD. We selected an arbitrary mass scale of $M=50M_{\odot}$ for all waveforms for demonstration, as this is a commonly observed mass scale in LIGO observations. 
%\te{This choice seems pretty arbitrary. Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70. This choice is important and should be more informed.}We will then examine the differences in optimization.

@@

We point out that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}. In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
waveforms while the rest of the waveforms are taken from NR simulations.
Instead, we solely use NR waveforms for comparison, since we are only exploring the possibility of optimizing waveform models. Thus, for simplicity sake, we ignore this procedure. On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
which they are long enough to contain part of the inspiral segment and all
merger and ringdown frequency information. We take the frequency limits as
$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
the IMRPhenomD's frequency range, except the minimum frequency is set higher
than that in the original calibration due to NR length. When compared with
IMRPhenomC, the frequency range is slightly extended to have a higher maximum
frequency. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
which is sufficient to capture all features of GW strain.

With the loss function evaluated, we apply gradient descent to optimize the tunable
coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
original coefficients given in \citep{khan2016frequency}. We take them as the initial 
waveform coefficients because they lie in the neighborhood of the minimum that we wish 
to find. Then, by taking $\alpha=10^{-6}$, we see the validation loss in 
Fig.~\ref{fig:loss} plateau at around $N=12000$, hence we end the optimization here.  
%\te{I think we need a bit more discussion about the stopping criterion here. What is N, and why did we choose it to be this number? Ideally we would actually have a plot of the loss function during training. I don't have a good intuition for if its noisy. Does it plateau? If we wanted to ensure that people don't complain about overfitting we could also plot the loss of the validation set to show its not going up.}

@@

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio
	$q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed
	here, 9 of them are also used in the original IMRPhenomD calibration.
	\citep{khan2016frequency} The two remaining waveforms were from BAM
	simulation, to which we do not have access.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.
